---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


## Overview

**tidyactuarial** is a comprehensive R package for financial mathematics and life contingencies. It bridges the gap between classical actuarial notation (SOA/CAS standards) and the modern **tidyverse** workflow.

The package is designed for actuaries, students, and researchers who need reproducible, vectorized, and "tidy" actuarial calculations.

## Key Features

  * **Life Contingencies:** Valuation of life insurance, annuities, and premiums for single and multiple lives.
  * **Flexible Assumptions:** Supports UDD (Uniform Distribution of Deaths), Constant Force, and Balducci assumptions for fractional ages.
  * **Financial Math:** Bond pricing (including callable bonds), yield curve interpolation (Spot/Forward rates), and immunization (Redington/Full).
  * **Tidy Design:** All main functions return `tibble` objects and are compatible with the native R pipe `|>` or `%>%`.

## Installation

Once the package is on CRAN, you can install it with:

```{r}
install.packages("tidyactuarial")
```

Alternatively, you can install the development version from [GitHub](https://www.google.com/search?q=https://github.com/jfajardo1908/tidyactuarial) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("julian.fajardo1908/tidyactuarial")
```

## Quick Start

### 1\. Life Tables and Premiums

Building a life table and calculating a net level premium for a 20-year endowment insurance:

```{r example_life}
library(tidyactuarial)

# 1. Create a life table (ensure qx has the same length as x)
ages <- 0:110
probs <- rep(0.002, length(ages)) 

lt <- lifetable(x = ages, qx = probs, radix = 100000)

# 2. Calculate net premium for a 35-year-old (20-year endowment)
# Added benefit = 100000
premium_x(
  lt = lt, 
  x = 35, 
  n = 20, 
  i = 0.05, 
  product = "endowment", 
  benefit = 100000,
  tidy = TRUE
)

```

### 2\. Bond Cash Flows

Generate a tidy schedule for a coupon bond:

```{r example_bond}
bond_cash_flows(
  face = 1000, 
  coupon_rate = 0.06, 
  years_to_maturity = 5, 
  coupons_per_year = 2
)
```

## References

Mathematical formulas and actuarial notation follow:

  * Finan, M. B. *A Reading of the Theory of Life Contingencies*.
  * Kellison, S. G. *The Theory of Interest*.
  * Dickson, D. C., Hardy, M. R., & Waters, H. R. *Actuarial Mathematics for Life Contingent Risks*.

